On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-29 2_Background.tex
fix some misleading information (related-work)

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}. Atoms of confusion is an important field of program comprehension introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.
%\castor{I'd rather have two concrete examples here, instead of this.}
They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.

@@

Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} also presented the results of a comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages. Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common. In addition, the imperative aspects of the language are syntactically similar to \clang, e.g., assignments, conditional expressions, \texttt{if}-statements, pre and post increments and decrements, among others. On the one hand, this means that some of the atoms of confusion that exist in \clang programs may also occur in JavaScript code. On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.
%Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm. Also, both languages share many syntactical rules, which allows us to replicate some of the C atoms in JavaScript.

@@

According to Arnaoudova et al.~\cite{Arnaoudova:2016:LAW}, Linguistic Antipatterns (LA) are \textit{``recurring poor practices in the naming, documentation, and choice of identifiers in the implementation of an entity [..] that may impair program understanding''}. These antipatterns are design issues that go against developer's  intuition about a program, for example, a \texttt{set} method that returns a result. Similarly to atoms of confusion, LAs occur at the source code level and hinder code readability. Differently from atoms of confusion, LAs are more akin to design principles than code idioms. In addition, they are language-agnostic, unlike atoms of confusion.

Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability. Some of them, such as avoiding more than three levels of code block nesting, are atom candidates. 
% Diego: Troquei "this work" para "their work". Removi o "more" de "are more related"
Their work evaluates the coding practices solely based on personal opinions of the study subjects. Furthermore, it focuses on the Java language. Finally, most of the practices analyzed in this study are more related to code formatting than language constructs or idioms.

>>>>

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}. Atoms of confusion is an important field of program comprehension introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. In a different work, Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} presented the results of a comprehensive research on the incidence of atoms of confusion in the wild, considering open-source projects written in \clang and \cpplang~programming languages.

@@

These previous studies~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17,DBLP:conf/msr/GopsteinZFC18}
have a greate influence in our work. Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common. In addition, the imperative aspects of the language are syntactically similar to \clang, e.g., assignments, conditional expressions, \texttt{if}-statements, pre and post increments and decrements, among others. On the one hand, this means that some of the atoms of confusion that exist in \clang programs may also occur in JavaScript code. On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.
%Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm. Also, both languages share many syntactical rules, which allows us to replicate some of the C atoms in JavaScript.

@@

%% According to Arnaoudova et al.~\cite{Arnaoudova:2016:LAW}, Linguistic Antipatterns (LA) are \textit{``recurring poor practices in the naming, documentation, and choice of identifiers in the implementation of an entity [..] that may impair program understanding''}. These antipatterns are design issues that go against developer's  intuition about a program, for example, a \texttt{set} method that returns a result. Similarly to atoms of confusion, LAs occur at the source code level and hinder code readability. Differently from atoms of confusion, LAs are more akin to design principles than code idioms. In addition, they are language-agnostic, unlike atoms of confusion.

Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability. Although some of these practices, such as avoiding more than three levels of code block nesting, are atom candidates, most of the practices analyzed in their study are more related to code formatting than language constructs or idioms. We borrow from their research the methodology for collecting the opinion of practitioners regarding code snippets preference.
